---
title: "Day 2"
draft: false
date: "2020-07-21T00:00:00+01:00"
output: html_document
menu:
  peer_network_school:
    parent: Daily Plan
    weight: 2
linktitle: "Day 2 - Individuals within networks"
toc: 
type: docs
weight: 1
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

If you missed it, [here is video of Workshop #2](<https://drexel.zoom.us/rec/share/99dVH-Guy09Jf7OV12PBAql6HJr6aaa81HUWqfUKyBn6z3EdwCdgLwjRhs_8OX9V>). 

Workshop #2 is going to build on the results of WS1, but we are actually going to analyze the network that we built in workshop #1. Also, I expect there will be a good deal more time available for discussion and questions. We will start with the network we made with attributes. We don't need to build the nework from scratch, instead I took the network we had and stored it as a .RDS file. RDS is the native structure for R.

In workshop #2 we will be looking at centrality metrics and network properties of individual nodes. For workshop #2 to work out ok, here are the things you will need to be able to do, which we covered in workshop #1.

- Install and load the following packages:
    - tidyverse
    - tidygraph
    - ggraph
    - igraph
    - here
- Read in an R object. 

In case you are interested, here are a few resources that I found useful for doing 
SNA:

# Reading (just some starting points)

- John Scott, and Peter J. Carrington, The SAGE Handbook of Social Network Analysis, 2014.
- Christina Prell, Social Network Analysis: History, Theory and Methodology, SAGE, 2011.
- Peter J. Carrington, John Scott, and Stanley Wasserman, Models and Methods in Social Network Analysis, Cambridge University Press, 2005.
- Stanley Wasserman and Katherine Faust, Social Network Analysis: Methods and Applications, Cambridge University Press, 1994.
- Mark E.J. Newman, Networks: An Introduction, Oxford University Press, 2010.
- Albert-Lazlo Barabasi, Network Science, freely available through Creative Commons license. http://barabasi.com/networksciencebook/
- Daniel Z. Grunspan, Benjamin L. Wiggins, and Steven M. Goodreau. (2014). ”Understanding classrooms through social network analysis: A primer for social network analysis in education research” CBE-Life Science Education 13(2) p. 167-178.http://www.lifescied.org/content/13/2/167.
- Katherine Ognyanova, (2015) ”Network Visualization with R”,  http://www.kateto.net/wordpress/wp-content/uploads/2015/06/Polnet%202015%20Network%20Viz%20Tutorial%20-%20Ognyanova.pdf
- Eric Brewe (2018) "The Roles of Engagement: Network Analysis in Physics Education Research" Getting Started in Physics Education Research https://www.compadre.org/Repository/document/ServeFile.cfm?ID=14725&DocID=4886


# Societies, Journals, and Meetings

## Professional Societies
- International Network of Social Network Analysis - Much more from the Social Network, social science side
- Network Science Society - The professional society that hosts NetSci - a premier network analysis conference, much more from the graph/theoretic world

## Journals
- Network Science - Cambridge Press Journal focused on Network Science - much more graph theoretic
- Connections - The journal of INSNA - much more focused on social network analysis & applications. 
- Social Networks - An Elsevier Journal focused on SNA - much more statistical

## Meetings
- Sunbelt - Conference of INSNA
- NetSci - Conference of Network Science Society
- CompleNet - International Conference on Complex Networks

# Datasets
- Colorado Index of Complex Networks - Maintained by Aaron Clauset - very high quality datasets
- Stanford Large Network Datasets - Maintained by Stanford - a list, not necessarily as clean, but a large list

The slides for Workshop #2 can be found here:
```{r WS1Slides}
knitr::include_url('https://ericbrewe.com/slides/rforsna/rforsna_ws2.html#1')
```

