---
title: "Day 2"
draft: false
date: "2020-07-21T00:00:00+01:00"
output: html_document
menu:
  peer_network_school:
    parent: Daily Plan
    weight: 2
linktitle: "Day 2 - Individuals within networks"
toc: 
type: docs
weight: 1
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```





Day #2 is going to build on the results of yesterday, but we are actually going to analyze the network that we built in workshop #1. Also, I expect there will be a good deal more time available for discussion and questions. 

The slides for Day #2 can be found here:
```{r Day2Slides}
knitr::include_url('https://ericbrewe.com/slides/PEER_network_school_2023/PEER_Individuals_Networks.html#1')
```

We will start with the network we made with attributes. We don't need to build the nework from scratch, instead I took the network we had and stored it as a .RDS file. RDS is the native structure for R.

In workshop #2 we will be looking at centrality metrics and network properties of individual nodes. For workshop #2 to work out ok, here are the things you will need to be able to do, which we covered in workshop #1.

- Install and load the following packages:
    - tidyverse
    - tidygraph
    - ggraph
    - igraph
    - here
- Read in an R object. 



---
Here are some readings on centrality as a set of connected metrics.


**Centrality metrics**
    - Borgatti, S. P. (2005). Centrality and network flow. Social networks, 27(1), 55-71. https://www.sciencedirect.com/science/article/abs/pii/S0378873304000693
    - Koschützki, D., Lehmann, K. A., Tenfelde-Podehl, D., & Zlotowski, O. (2005). Advanced centrality concepts. In Network analysis (pp. 83-111). Springer, Berlin, Heidelberg. https://link.springer.com/chapter/10.1007/978-3-540-31955-9_5
    
This one sort of spans centrality and missing data:
- Borgatti, S. P., Carley, K. M., & Krackhardt, D. (2006). On the robustness of centrality measures under conditions of imperfect data. Social networks, 28(2), 124-136. https://www.sciencedirect.com/science/article/abs/pii/S0378873305000353


