---
title: "Workshop 3"
draft: false
date: "2020-07-21T00:00:00+01:00"
output: html_document
menu:
  RForSNA:
    parent: Workshops
    weight: 3
linktitle: "Workshop 3"
toc: 
type: docs
weight: 1
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

Workshop #3 continues to build on the results of WS1, but we are actually going to analyze the network that we built in workshop #1. 

In workshop #3 we will be looking at characterizing the network as a whole, and thinking about how we can compare networks. 

There were a couple of questions that arose out of workshop #2 (one about missing data and one about choosing a centrality metric). I promised a couple of papers, and here they are:

**Missing Data**
    - Žnidaršič, A., Ferligoj, A., & Doreian, P. (2012). Non-response in social networks: The impact of different non-response treatments on the stability of blockmodels. Social Networks, 34(4), 438-450. https://www.sciencedirect.com/science/article/abs/pii/S037887331200010X
    - Huisman, M. (2009). Imputation of missing network data: Some simple procedures. Journal of Social Structure, 10(1), 1-29. https://www.rug.nl/research/portal/files/14031886/Huisman_imputation_JoSS2009.pdf
    
**Centrality metrics**
    - Borgatti, S. P. (2005). Centrality and network flow. Social networks, 27(1), 55-71. https://www.sciencedirect.com/science/article/abs/pii/S0378873304000693
    - Koschützki, D., Lehmann, K. A., Tenfelde-Podehl, D., & Zlotowski, O. (2005). Advanced centrality concepts. In Network analysis (pp. 83-111). Springer, Berlin, Heidelberg. https://link.springer.com/chapter/10.1007/978-3-540-31955-9_5
    
This one sort of spans the two:
- Borgatti, S. P., Carley, K. M., & Krackhardt, D. (2006). On the robustness of centrality measures under conditions of imperfect data. Social networks, 28(2), 124-136. https://www.sciencedirect.com/science/article/abs/pii/S0378873305000353



The slides for Workshop #3 can be found here:
```{r WS2Slides}
knitr::include_url('https://ericbrewe.com/slides/rforsna/rforsna_ws3.html#1')
```

