---
title: "R For SNA"
subtitle: Workshop 2
author: 'Eric Brewe <br> Associate Professor of Physics at Drexel University <br>'
date: "11 August 2020, last update: `r Sys.Date()`"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, "BreweSlides.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse) # load tidyverse package
library(emo) # load emoji package
library(igraph)
library(tidygraph)
library(ggraph)
library(here)

```

---
# Welcome Back!

### Workshop #1 was a sprint!
This should be a bit more reasonably paced. 

First of all, we will not need to build the network all over again. 

But we should start with some [project-oriented workflow.](<https://www.tidyverse.org/blog/2017/12/workflow-vs-script/>)

The blog post by Jenny Bryan linked above describes the basic logic for a project orinted workflow. 

Each project gets:
- It's own folder
- A project file (which you never open)
- As many Rmarkdown or R script files as needed
- Subfolders for:
    - data
    - plots
    - reports/papers etc.

---
# Let's build on our project oriented workflow




```{r LoadPackages, eval=FALSE}
library(tidyverse) #tools for cleaning data 
library(igraph)  #package for doing network analysis
library(tidygraph) #tools for doing tidy networks
library(here) #tools for project-based workflow
library(ggraph) #plotting tools for networks
```


