---
title: "PEER Advanced Field School 2023"
subtitle: Day 1 - Theory, Data, and Plotting
author: 'Eric Brewe <br> Professor of Physics at Drexel University <br>'
date: "15 June 2023, last update: `r Sys.Date()`"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, "BreweSlides.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse) # load tidyverse package
library(emoji) # load emoji package
library(janitor)
library(glmnet)
library(haven)
library(here)

```




class: center, middle

# Introduction to R and plotting recipes

---

# What you need to know about R?

- R is the programming language.
- RStudio is the Integrated Development Environment (IDE)
- Posit.cloud is an online version of the IDE
- Packages: groups of functions that are developed as open source 
- Base R
    - The group of packages preloaded into R
- Tidyverse
    - Family of packages that are designed with the theory that programming should be readable by humans.


.font190[.center[Lets do this!]]

---

# How do we start learning R?

- We need to know something about the different types of data and the ways in which they are stored. 

.pull-left[

#### Data types (at least some of them)

- **Logical** (T/F, 1/0)
- **Integers** (whole numbers)
- **Numeric** (numbers with decimal places)
- **Complex** (I never use these)
- **Text** (exactly what it sounds like)
]

.pull-right[
#### Data storage (at least some of them)

- **Vectors** = long columns of data (can be any type, but only one type of data)
- **Dataframes** = like Excel pages (columns can hold different types of data)
- **Matrices** = like dataframes, but they have named rows/columns 
    - Adjacency Matrices are of type matrix.
- **Lists** = the junkdrawer, can hold any type of data (including dataframes or matrices)
    - in igraph, networks are stored as lists.

]
---


# Lets Take a Tour of RStudio IDE

---

# Let's Install Some Packages
.font150[You'll only need to do this once.]

## In Console

To install tidyverse package...

```{r InstallPackages, eval=FALSE}
install.packages("tidyverse")
```

Repeat this with the following packages:

glmnet janitor haven

---

# Let's Load Some Packages
.font150[You'll need to do this every time you restart R.]

## In Console

To load tidyverse package...

```{r LoadPackages, eval=FALSE}
library(tidyverse) #tools for cleaning and visualizing data 
library(janitor)  #package for cleaning data
library(glmnet) #tools for doing regression analyses
library(here) #tools for project-based workflow
library(haven) #tools for reading in data from spss
```

--

Once you have done this, you will want to put include a code chunk with all of your libraries into your markdown document so that you don't have to type this every time. 

---
# Let's get data into R. 

In this workshop, you have the data from the survey to make things as easy as possible. However, this is not typical.  You will have to find ways to get your data into R but that is sort of different depending on where you are running R (posit.cloud or a local machine)

If you have loaded the package "here" this should just work.  If you have not loaded the "here" package you will need to set the working directory. 

Again, you will want to include this as a code chunk in your RMD file.  



```{r ReadDataforReal, echo=FALSE}
#This loads the csv and saves it as a dataframe titled week_1_data

week_1_data <- read_csv(here("static/slides/EDUC_847/data", "EDUC_847_Survey.csv"))

```


```{r ReadData, eval=FALSE}
#This loads the csv and saves it as a dataframe titled week_1_data

week_1_data <- read_csv(here("static/slides/EDUC_847/data", "EDUC_847_Survey.csv"))

```


---
# Let's have a look at the data

```{r GlimpseData}
glimpse(week_1_data)

```



---
# Let's start cleaning up.

## First, we don't need most of that data

There is a ton of data there that doesn't make sense for us to keep around. 

We will use the '%>%' (pipe) operator and the verb select

```{r DeleteUnNeededVariables}
week_1_data %>%
  select(Q2:Q10) -> w1df #<<

glimpse(w1df)
```

???
Note the data are not numbers.
---
# Let's keep cleaning up.

Notice there are two rows of data that are just the questions and something from Quqltrics.  Let's get rid of those. 

```{r RemoveExtraRows}
w1df %>%
  filter(row_number() >3 ) -> w1df

```


---
# Let's check out the power of R

.font200[.center[I am going to blast through these next slides, to show you some of the things that you might want to do with R]]


---
# Let's start visualizing the data

## For categorical data, you might want to get some counts.

Here is code to do this for the question about morning or night person.


```{r MorningOrNight}

w1df %>%
  select(Q2) %>%  #<<
  group_by(Q2) %>% #<<
  tally() 


```


---
# Let's start visualizing the data

## For categorical data, you might want to get a histogram.

Histograms are very important for regression!

Here is code to do this for the favorite dessert type.

.pull-left[
```{r DessertType}

w1df %>%
  select(Q3) %>%
  group_by(Q3) %>% 
  tally()  

```
]

.pull-right[
```{r DessertTypeHist, out.width= "50%"}

w1df %>%
  select(Q3) %>%
  ggplot(aes(y = Q3)) + 
  geom_bar()

```
]

---
# Let's explore beverage choices

```{r BevChoices}
w1df %>%
  select(Q4) %>%
  head()
```

.font150[Notice, these are *not* tidy data, more than one variable per line]
---
# Let's dummy code the beverage data
This is going to be useful later in the term I promise!
.pull-left[
### For **tidy** data we want one value per row

```{r TidyBevChoices}
w1df %>%
  select(Q4) %>%
  separate_rows(Q4, sep = ",") %>% #<<
  head(10)
```

]

.pull-right[
### We can dummy code these

```{r WideTidyBevChoices}
WorkshopData %>%
  select(ID, Q4) %>%
  separate_rows(Q4, sep = ",") %>%
  mutate(Checked = 1) %>% #<<
  pivot_wider(names_from = Q4, #<<
              values_from = Checked, #<<
              values_fill = 0) #<<

```


]


---
# Let's look at some quantitative data

### First, let's summarize the reading data.

```{r SummarizeBook,}

WorkshopData %>%
  select(Q5) %>%
  summarize(Ave = mean(Q5, na.rm = TRUE), 
            SD = sd(Q5, na.rm = TRUE))

```


---
# Let's investigate groups

.pull-left[
### Are morning people or night owls reading longer books?

```{r ReadersTimeOfDay}

WorkshopData %>%
  select(Q2, Q5) %>%
  group_by(Q2) %>%
  summarize(Ave = mean(Q5), #<<
            SD = sd(Q5)) #<<

```
]
.pull-right[

### We might want to use a boxplot to display these data

```{r ReadersTimeOfDayBoxplot, out.width = "70%"}

WorkshopData %>%
  select(Q2, Q5) %>%
  ggplot(., aes(x = Q2, y = Q5)) +
  geom_boxplot()

```
]

---
# Let's look at readers vs. blueberries

### Is there a relationship between length of book and estimates on number of blueberries?

.pull-left[
### Could do a scatter plot
```{r ScatterPlot, fig.show= 'hide'}

WorkshopData %>%
  select(Q5:Q6) %>%
  mutate(Q5 = as.numeric(Q5), Q6 = as.numeric(Q6)) %>%
  ggplot(aes(x = Q5, y = Q6)) +
  geom_point()

```

]
.pull-right[
```{r ref.label= 'ScatterPlot', echo=FALSE}

```

]

---
# Let's look at readers vs. blueberries

### Is there a relationship between length of book and estimates on number of blueberries?

### Or you could do a linear model

```{r LinearModel}

summary(lm(Q6 ~ Q5, data = WorkshopData))

```
---
# Let's prep our data for SNA

### We will need to prep two separate files...

1. An edgelist
2. A file of attributes of the nodes. 

---
# Let's make an edgelist


There is an issue here. 

I wanted to make these data anonymous (so we don't know who likes scotch)

But to do that I had to make the edgelist for you. 

So I sent you an edgelist as a csv. Sorry.


### Check out the edgelist

```{r GenerateEL, eval=FALSE}
EL = read_csv(here("data", "AnonEL.csv"))
head(EL)
```

```{r GenerateELForReal, echo=FALSE}
EL = read_csv(here("static/slides/PEER_network_school_2023/data", "AnonEL.csv"))
head(EL)
```



---
# Let's assemble our node attributes

Before we can convert our Edgelist to a network, we should add in the attributes. 

We have several candidate attributes:
- Morning vs. Night
- Dessert Type
- Pages in book

We will develop a separate dataframe for the attributes. 

```{r Attributes}
WorkshopData %>%
  select(ID, Q2, Q3, Q5) -> AttributeDf
```


---
# Let's assemble our node attributes

Experience tells me that when you try to add attributes that you often make a mistake where the number attributes don't match up well to the number of nodes...but lets see.




```{r AssembleGraph, out.width= "40%"}

gr <- graph_from_data_frame(EL, directed = TRUE)

plot(gr)


gr = as_tbl_graph(gr)

```



---
# Let's add some attributes

So actually the easiest way to add the attributes is to add them while you make the graph.  

But that isn't as easy as it seems

```{r GraphWAttributes1, warning=TRUE, eval=FALSE}

gr %>%
  activate(nodes) %>%
  mutate(AMPM = AttributeDf$Q2)

```
But that isn't as easy as it seems...


--- 
# Let's sort out these attributes. 



The warning was: "Input `AMPM` must be size 60 or 1, not 34."

What this means is we need to take our attributes dataframe and make sure all the nodes are listed. 

To do this we need to:
- Compile a dataframe of all nodes listed in the graph
- Use a join to add attributes to this dataframe

---
# Let's sort out these attributes (pt 2)

```{r GraphWAttributes2}
#This will get a vector of all nodes
gr %>%
  activate(nodes) %>%
  as_tibble() %>%
  transmute(ID = name) %>%
  mutate(ID = as.numeric(ID))-> GrNodes

#Now we pull in the attributes using a left_join
NodeAttributes = left_join(GrNodes, AttributeDf, by = "ID")
```

You should inspect Node Attributes

---
# Let's sort out these attributes (pt 3)


### Inspect the head

```{r GraphWAttributes3}

head(NodeAttributes)

```

---
# Let's sort out these attributes (pt 4)

### Inspect the tail

```{r GraphWAttributes4}

tail(NodeAttributes)

```



---
# Let's finally assemble this graph. 

```{r BuildGraph}

gr %>%
  as_tbl_graph() %>%
  activate(nodes) %>%
  mutate(AMPM = NodeAttributes$Q2) %>%
  mutate(Dessert = NodeAttributes$Q3) %>%
  mutate(Pages = NodeAttributes$Q5) -> gr

summary(gr)
```
---
# Let's finally assemble this graph.

### Take a look at your graph

```{r InspectGr, echo=FALSE}
 gr 
```

---
# Let's make our plots look a bit better
.pull-left[
### We can add various elements to our plot
- Color (good for grouping)
- Shape (good for grouping)
- Linewidth (good for numeric)
- Text (good for labels)
- Layout

### This is the most basic plot
````
```{r PlotGr1, eval = FALSE}
ggraph(gr) +
  geom_edge_link() +
  geom_node_point()
```
````
]

.pull-right[

```{r PlotGr1a, echo=FALSE}
ggraph(gr) +
  geom_edge_link() +
  geom_node_point()
```

Not super pretty
]

---
# Let's make our plots look a bit better
.pull-left[
### We can add various elements to our plot
- Layout, there are lots of options
    -Try circle?

````
```{r PlotGr2, eval = F}
ggraph(gr, layout = 'circle') +
  geom_edge_link() +
  geom_node_point()
```
````
]

.pull-right[
```{r PlotGr2a, echo = F}
ggraph(gr, layout = 'circle') +
  geom_edge_link() +
  geom_node_point()
```
]

---

# Let's make our plots look a bit better

.pull-left[
### We can add various elements to our plot
- Shape, we can make night owls one shape and morning people a differnt shape...

````
```{r PlotGr3, eval = F}
ggraph(gr, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(shape = AMPM))
```
````
]
.pull-right[
```{r PlotGr3a, echo = F}
ggraph(gr, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(shape = AMPM))
```

### Not great

]
---

# Let's make our plots look a bit better

.pull-left[
### We can add various elements to our plot
- Color, lets use the dessert type to define a color

````
```{r PlotGr4,eval=FALSE}
ggraph(gr, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(color = Dessert))
```
````
]
.pull-right[
```{r PlotGr4a, echo=FALSE}
ggraph(gr, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(color = Dessert))
```
]
---

# Let's make our plots look a bit better
.pull-left[
### We can add various elements to our plot
- Size, lets make the nodes different sizes based on the number of pages in the last book read.

````
```{r PlotGr5, eval=FALSE}
ggraph(gr, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(size = Pages))
```
````
]
.pull-right[
```{r PlotGr5a, echo=FALSE}
ggraph(gr, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(size = Pages))
```
]
---

# Let's make our plots look a bit better

.pull-left[
### We can add various elements to our plot
- Let's jam them all together!

````
```{r PlotGr6, eval=FALSE}
ggraph(gr, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(shape = AMPM, 
                      color = Dessert, 
                      size = Pages))
```
````
]

.pull-right[

```{r PlotGr6a, echo=FALSE}
ggraph(gr, layout = 'circle') +
  geom_edge_link() +
  geom_node_point(aes(shape = AMPM, 
                      color = Dessert, 
                      size = Pages))
```

]

---

# Let's save our data so that we can access it easily

.pull-left[
### We want to save as an RDS file
- This way we avoid having to reconstruct our graph
- We get our nodes, our edges, and our attributes.

You don't actually need to run this (I've already saved the file as an RDS so we can start with that on Day 2!)


]

.pull-right[

````
```{r WriteRDS, eval=FALSE}
write_rds(gr, here("data", "WorkshopNetwork.rds"))
```
````
]