---
title: "EDUC 847 Winter 24"
subtitle: Week 2 - Fundamentals of Linear Modeling
author: 'Eric Brewe <br> Professor of Physics at Drexel University <br>'
date: "22 January 2024, last update: `r Sys.Date()`"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, "BreweSlides.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse) # load tidyverse package
library(emoji) # load emoji package
library(janitor)
library(glmnet)
library(haven)
library(here)
library(broom)

```




class: center, middle

# Let's start where we ended last class...Scatter Plots

Start by getting the data into R

```{r ReadDataforReal, echo=FALSE}

week_1_data <- read_csv(here("static/slides/EDUC_847/data", "EDUC_847_Survey.csv"))

```


```{r ReadData, eval=FALSE}
#This loads the csv and saves it as a dataframe titled week_1_data

week_1_data <- read_csv(here("static/slides/EDUC_847/data",
                             "EDUC_847_Survey.csv"))

```


---
# Let's start cleaning up.

## First, we don't need most of that data

There is a ton of data there that doesn't make sense for us to keep around. 

We will use the '%>%' (pipe) operator and the verb select

```{r DeleteUnNeededVariables}
week_1_data %>%
  select(ID,Q2:Q6) -> w1df #<<

glimpse(w1df)
```


---
# Let's keep cleaning up.

Notice there are two rows of data that are just the questions and something from Qualtrics.  Let's get rid of those. 

```{r RemoveExtraRows}
w1df %>%
  filter(row_number() >3 ) -> w1df

```

---
# Let's lump all this together so we don't have to do it again.


```{r CleanDataPipeline}

week_1_data %>%
  select(ID,Q2:Q6) %>%
  filter(row_number() >3 ) %>%
  mutate(across(c(Q2:Q3,Q5:Q6), as.numeric)) -> w1df  #<<
  

```


---
# Let's look at books vs. weight of cow

### Is there a relationship between length of book and estimates on weight of cow?

.pull-left[
### Making a scatter plot to explore
```{r ScatterPlot2, fig.show= 'hide'}

w1df %>%
  select(Q5:Q6) %>%
  ggplot(aes(x = Q5, y = Q6)) +
  geom_point() + 
  theme_minimal() + 
  labs(x = "Pages", 
       y = "Weight of Cow",
       title = "Kick Ass Scatterplot!")

```

]

.pull-right[
```{r ref.label= 'ScatterPlot2', echo=FALSE}

```

]


---

```{r ScatterPlotWithResiduals, fig.show= 'hide'}
mod <- lm(Q5 ~ Q6, data = w1df)
summary(mod)
mod_df<-augment(mod)


mod_df %>%
  ggplot(aes(x = Q6, y = Q5)) +
  geom_point() + 
  theme_minimal() + 
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = Q6, yend= .fitted), color = "red", size = 0.2) +
  labs(x = "Pages", 
       y = "Weight of Cow",
       title = "Kick Ass Scatterplot!")

```



---
# Let's make the histogram even better


.pull-left[
```{r AMPM5}

w1df %>%
  select(Q2) %>%  #<<
  group_by(Q2) %>% #<<
  tally() 


```
]

.pull-right[
```{r AMPMHist4, out.width= "50%"}

w1df %>%
  select(Q2) %>%
  ggplot(aes(y = Q2)) + 
  geom_bar() + 
  theme_minimal() + #<<
  ggtitle("Histogram of morning/night") #<<


```
]

Good enough for now...you are always welcome to continue adding layers to your plot to make it look even better.  
See here for more: https://ggplot2.tidyverse.org/





---
# Let's use some colors

What if Morning People have different relationships between the estimates of cow weight and length of the closest book.

.pull-left[
### Could do a scatter plot
```{r ScatterPlot3, fig.show= 'hide'}

w1df %>%
  select(Q2,Q5:Q6) %>%
  ggplot(aes(x = Q5, y = Q6, color = Q2)) +
  geom_point() + 
  theme_minimal() + 
  labs(x = "Pages", 
       y = "Weight of Cow",
       title = "Scatterplot!")

```

]

.pull-right[
```{r ref.label= 'ScatterPlot3', echo=FALSE}

```

]

---
# Let's recap

- We know how to get data into R
- We know how to modify the data
- We know how to plot the data
- We have recipes for plotting 
    - histograms
    - boxplots
    - scatterplots
    
We are ready to go!
    