---
title: "Week 1"
draft: false
date: "2020-07-21T00:00:00+01:00"
output: html_document
menu:
  applied_regression:
    parent: Weekly Plan
    weight: 1
linktitle: "Week 1 - Theory and Data"
toc: 
type: docs
weight: 1
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```




The slides for Week #1 can be found here:
```{r WS1Slides}
knitr::include_url('https://ericbrewe.com/slides/EDUC_847/EDUC_847_Week_1.html#1')
```



**Reading Assignment**
Here is your reading assignment for Week 1

Read Chapter 5 in The Effect: Research Design and Causality  https://theeffectbook.net/ch-Identification.html

**Additional Resources**
If you are struggling with R, I think that the introduction chapters of *Learning Statistics with R* are really useful, I suggest looking at chapters 3-8: https://learningstatisticswithr.com/book/part-ii-an-introduction-to-r.html 
[WTF R](<https://rstats.wtf/index.html>) Jenny Bryan is great at explaining things about R and so frequently, I find myself looking at her books.