---
title: "Day 1"
draft: false
date: "2020-07-21T00:00:00+01:00"
output: html_document
menu:
  peer_network_school:
    parent: Daily Plan
    weight: 1
linktitle: "Day 1 - Theory and Data"
toc: 
type: docs
weight: 1
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```




The slides for Day #1 can be found here:
```{r WS1Slides}
knitr::include_url('https://ericbrewe.com/slides/PEER_network_school_2023/PEER_network_theory.html#1')
```


Each day I am going to include some extra reading materials and notes in case you want to go further.  For day 1 here are some readings on missing data and network generators

**Missing Data**
    - Žnidaršič, A., Ferligoj, A., & Doreian, P. (2012). Non-response in social networks: The impact of different non-response treatments on the stability of blockmodels. Social Networks, 34(4), 438-450. https://www.sciencedirect.com/science/article/abs/pii/S037887331200010X
    - Huisman, M. (2009). Imputation of missing network data: Some simple procedures. Journal of Social Structure, 10(1), 1-29. https://www.rug.nl/research/portal/files/14031886/Huisman_imputation_JoSS2009.pdf
