---
title: "Week 4"
draft: false
date: "2020-07-21T00:00:00+01:00"
output: html_document
menu:
  applied_regression:
    parent: Weekly Plan
    weight: 1
linktitle: "Week 4 - Data for Linear Modeling"
toc: 
type: docs
weight: 1
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
library(tidyverse)
```




The slides for Week 4 can be found here:
```{r WS1Slides}
knitr::include_url('https://ericbrewe.com/slides/EDUC_847/EDUC_847_Week_4.html#1')
```



**Reading Assignment**
Here is your reading assignment for Week 4

Read  Azur, M. J., Stuart, E. A., Frangakis, C., & Leaf, P. J. (2011). Multiple imputation by chained equations: what is it and how does it work?. International journal of methods in psychiatric research, 20(1), 40-49
https://onlinelibrary.wiley.com/doi/10.1002/mpr.329


**Additional Resources**
If you are struggling with R, I think that the introduction chapters of *Learning Statistics with R* are really useful, I suggest looking at chapters 3-8: https://learningstatisticswithr.com/book/part-ii-an-introduction-to-r.html 
