---
title: "PEER Advanced Field School 2023"
subtitle: Day 3 - Whole Networks
author: 'Eric Brewe <br> Professor of Physics at Drexel University <br>'
date: "15 June 2023, last update: `r Sys.Date()`"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, "BreweSlides.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse) # load tidyverse package
library(emoji) # load emoji package
library(igraph)
library(tidygraph)
library(ggraph)
library(here)
library(boot)
```


# Welcome Back!

### Our project is about building and analyzing the Workshop network

We will continue to use the data file from yesterday, WorkshopNetwork.rds. 


```{r LoadPackages, eval=FALSE}
library(tidyverse) #tools for cleaning data 
library(igraph)  #package for doing network analysis
library(tidygraph) #tools for doing tidy networks
library(here) #tools for project-based workflow
library(ggraph) #plotting tools for networks
library(boot)  #to do resampling
```


---
# Let's load our data 



```{r LoadNetwork, eval=FALSE}

gr <- readRDS(here("data", "WorkshopNetwork.rds"))

```

```{r LoadNetworkForReal, echo=FALSE}

gr <- readRDS(here("static/slides/PEER_network_school_2023/data", "WorkshopNetwork.rds"))

```

---
# Let's plot it quickly

While we are at it we might as well set a layout.

.pull-left[
## Whare are distinctive features of the network?

````
```{r plotGr, eval = F}
GrLayout <- create_layout(gr,
                          layout = "kk")

ggraph(GrLayout) +
  geom_edge_link() +
  geom_node_point() +
  theme(legend.position="bottom")
```
````

]

.pull-right[
```{r plotGrR, echo = F}
GrLayout <- create_layout(gr,
                          layout = "kk")

ggraph(GrLayout) +
  geom_edge_link() +
  geom_node_point() +
  theme(legend.position="bottom")
```
]
---
# Let's explore whole network metrics

**These are fairly easy to calculate**
- Density: Proportion of possible connections that exist
- Diameter: Longest Shortest Path across network
- Reciprocity: Extent to which each directed edge is a two way edge
- Transitivity: Extent to which three connected nodes form triangles
- Average Path Length: Average shortest distance between every pair of nodes

**These require a bit more work**
- Giant: Number of nodes connected to largest component
- Homophily: Extent to which nodes with similar attribute are connected
- Average degree: This is pretty descriptive. 


This is a list of many common graph metrics used in tidygraph
https://rdrr.io/cran/tidygraph/man/graph_measures.html

---
# Lets calculate a couple of these...

### For many of these it is a very simple one line command.
Here is density
```{r CalcDensity}

edge_density(gr)

```

Here is diameter, note that with tidygraph you have to use the with_graph() function, and specify the graph and function graph_diameter()
```{r CalcDiameter}

with_graph(gr, graph_diameter())

```

--
.center.font200[So?]


---
# Let's calculate a couple of others

###  It is your turn, choose one of the ones that are easy to calculate and make the calculation. 

--
```{r}

with_graph(gr, graph_reciprocity()) #Reciprocity
transitivity(gr) #Transitivity
with_graph(gr, graph_mean_dist()) #Average Distance

```

---
# Let's create a function

### Creating a function accomplishes two things:
1. We can calculate some of the harder to calcluate metrics
2. We do not have to re-do each time we want to analyze a new network.

```{r GrMetricFunction}
GrFunction <- function(gr){
  Giant = max(clusters(gr)$csize)
  AveDeg = gr %>%
    activate(nodes) %>%
    mutate(Deg = centrality_degree( mode = 'total') ) %>%
    select(Deg) %>%
    as_tibble() %>%
    summarise(AveDeg = mean(Deg))
  df = tibble(Giant, AveDeg )
  return(df)
}

```

---
# Let's run this function

```{r RunGrFunction}

GrFunction(gr)
```

### What do these numbers mean?
--

*Giant* = 30, there are 30 people connected in the largest component. 

*AveDeg*, the average person has 3.37 incoming or outgoing edges

---

# Let's talk about clustering

Whole networks sometimes have regions which are different. 

Perhaps you want to know, are there groups that have some common characteristic?

So we need a way to look for groups within our network.  

### Start with random assignment:


````
```{r RandGrGr}
#First generate a grouping based on random assignment.
gr %>%
  activate(nodes) %>%
  mutate(RandGroup = sample(1:2,60, replace = TRUE)) -> gr

```
````

---

# Let's plot this with the random assignment to groups

.pull-left[

````
```{r PlotRandGroupGr, eval = F}
GrLayout <- create_layout(gr,
                          layout = "kk")

ggraph(GrLayout) +
  geom_edge_link() +
  geom_node_point(aes(color = factor(RandGroup))) + #<<
  theme(legend.position="bottom")
```
````

]

.pull-right[


```{r PlotRandGroupGrR, echo = F}
GrLayout <- create_layout(gr,
                          layout = "kk")

ggraph(GrLayout) +
  geom_edge_link() +
  geom_node_point(aes(color = factor(RandGroup))) +
  theme(legend.position="bottom")
```


]
---
# Let's talk about Modularity

Modularity is a metric that tells us whether there is a way to break up our network into chunks which have more connections within the chunk than from the chunk to other parts of the network. 

```{r RandGrMod}
with_graph(gr, graph_modularity(group = RandGroup)) #Modularity
```

So the modularity of our network with nodes randomly assigned to groups is .12

This means that based on our grouping, there is a very slightly greater proportion of links between members of the group than across groups. 




---
# Let's use an actual community detection algorithm

We'll use Walktrap, in which a random surfer can jump node to node based on the presence of edges.  Then the walktrap algorithm establishes community structure by finding the propensity to end up in a cluster of nodes. 

There are two ways to do this...
```{r WalkTrapGr}

#First generate a grouping based on the walktrap algorithm and add it to the graph.
gr %>%
  activate(nodes) %>%
  mutate(WTGroup = group_walktrap()) -> gr

#The other way is to just call the walktrap algorithm. 
wtg <- with_graph(gr, group_walktrap())
wtg
```

---
# Let's talk about Modularity

Now we can check the modularity with the walktrap defined communities

```{r WTGrMod}
with_graph(gr, graph_modularity(group = WTGroup)) #Modularity
```

So the modularity of our network with nodes assigned to groups via the walktrap algorithm is .74

This means that based on our grouping, there is a much greater proportion of links between members of the group than across groups. 

---
# Let's plot the walktrap communities

This is the hard way to plot these.

.pull-left[

````
```{r PlotWTGroupGr, eval = F}
GrLayout <- create_layout(gr,
                          layout = "kk")

ggraph(GrLayout) +
  geom_edge_link() +
  geom_node_point(aes(color = factor(WTGroup))) + #<<
  theme(legend.position="bottom")
```
````

]

.pull-right[


```{r PlotWTGroupGrR, echo = F}
GrLayout <- create_layout(gr,
                          layout = "kk")

ggraph(GrLayout) +
  geom_edge_link() +
  geom_node_point(aes(color = factor(WTGroup))) +
  theme(legend.position="bottom")
```


]

---
# Let's plot the walktrap communities again

This is the easy way to plot these (though not tidy).

.pull-left[

````
```{r PlotWTGroupEasyGr, eval = F}

wtg = cluster_walktrap(gr)

modularity(wtg)

plot(wtg, gr)
```
````

]

.pull-right[


```{r PlotWTGroupEasyGrR, echo = F}

wtg = cluster_walktrap(gr)

modularity(wtg)

plot(wtg, gr)
```

]
---
# Now it is your turn

- Choose a community detection algorithm, 
- calculate it for the Workshop Network, and 
- be ready to talk about what the salient points of the community detection algorithm you chose. 

Here are the community detection algorithms available to you in tidygraph.
https://tidygraph.data-imaginist.com/reference/index.html#community-detection
