---
title: "EDUC 847 Winter 24"
subtitle: Week 6 - Logistic Regression
author: 'Eric Brewe <br> Professor of Physics at Drexel University <br>'
date: "26 February 2024, last update: `r Sys.Date()`"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, "BreweSlides.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(haven)
library(here)
library(broom)
library(mice)
library(car)
library(jtools)
library(kableExtra)
library(broom.mixed)
library(emdi)
library(interactions)

```





# Let's start with a complete dataset

Start by getting the data into R

```{r ReadDataforReal, echo=FALSE}

sci_df <- read_csv(here("static/slides/EDUC_847/data", "sci_scores.csv"))

```


```{r ReadData, eval=FALSE}
#This loads the csv and saves it as a dataframe titled sci_df

sci_df <- read_csv(here("data",
                        "sci_scores.csv"))

```


---
# Let's remember last class

### We predicted science scores from math scores. 
```{r LinearRegression}

sci_mod_math <- lm(sci ~ math_sc, data = sci_df) #<<

tidy(summary(sci_mod_math))
```

Note - we have a new dataset, nothing is missing!

---
# Let's go futher

## What if we wanted to then predict who would pass some class. 

### Probably need to look at our data


```{r ReviewData}

glimpse(sci_df)

```

---
# Let's plot 

## We can try a scatter plot

```{r scatterplot}
sci_df %>%
  ggplot(aes(x = math_sc, y = pass)) +
  geom_point()

```


---
# Let's plot 

## Boxplots should be better

```{r boxplot}
sci_df %>%
  ggplot(aes(x = as.factor(pass), y = math_sc)) +
  geom_boxplot()

```

---
# Let's predict passing

## Here's what we know:

- Pass / Fail is a binary outcome variable
- We have a variety of types of predictor variables



---
# Let's review...

We spent today looking at how to 

We can...
 - run a multiple regression
 - interpret a multiple regression 
 - consider categorical variables
 - dummy code for categorical variables
 - visualize regression coefficients
 - handle interaction terms