---
title: "EDUC 847 Winter 25"
subtitle: Week 2 - Fundamentals of Linear Modeling
author: 'Eric Brewe <br> Professor of Physics at Drexel University <br>'
date: "15 January 2025, last update: `r Sys.Date()`"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, "BreweSlides.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse) # load tidyverse package
library(emoji) # load emoji package
library(janitor)
library(glmnet)
library(haven)
library(here)
library(broom)

```




class: center, middle

# Let's start where we ended last class...Scatter Plots

Start by getting the data into R

```{r ReadDataforReal, echo=FALSE}

week_1_data <- read_csv(here("static/slides/EDUC_847/data", "EDUC_847_Survey.csv"))

```


```{r ReadData, eval=FALSE}
#This loads the csv and saves it as a dataframe titled week_1_data

week_1_data <- read_csv(here("static/slides/EDUC_847/data",
                             "EDUC_847_Survey.csv"))

```


---
# Let's start cleaning up.

## First, we don't need most of that data

There is a ton of data there that doesn't make sense for us to keep around. 

We will use the '%>%' (pipe) operator and the verb select

```{r DeleteUnNeededVariables}
week_1_data %>%
  select(ID,Q2:Q6) -> w1df #<<

glimpse(w1df)
```


---
# Let's keep cleaning up.

Notice there are two rows of data that are just the questions and something from Qualtrics.  Let's get rid of those. 

```{r RemoveExtraRows}
w1df %>%
  filter(row_number() >3 ) -> w1df

```

---
# Let's lump all this together so we don't have to do it again.


```{r CleanDataPipeline}

week_1_data %>%
  select(ID,Q2:Q6) %>%
  filter(row_number() >3 ) %>%
  mutate(across(c(Q2:Q3,Q5:Q6), as.numeric)) -> w1df  #<<
  

```


---
# Let's look at books vs. weight of cow

### Is there a relationship between length of book (Q5) and estimates on weight of cow (Q6)?

.pull-left[
### Making a scatter plot to explore
```{r ScatterPlot2, fig.show= 'hide'}

w1df %>%
  select(Q5:Q6) %>%
  ggplot(aes(x = Q5, y = Q6)) +
  geom_point() + 
  theme_minimal() + 
  labs(x = "Pages", 
       y = "Weight of Cow",
       title = "Kick Ass Scatterplot!")

```

]

.pull-right[
```{r ref.label= 'ScatterPlot2', echo=FALSE}

```

]

A **residual** is the distance between the Y value for each measurement, and the predicted Y value. 

---
# Let's investigate the residuals for our data 
.pull-left[
```{r CheckResiduals1}

w1df %>% 
  select(Q5,Q6)
```
]
.pull-right[
```{r ref.label= 'ScatterPlot2', echo=FALSE}

```
]

---
# Let's make a linear regression model


### We want the estimated weight of the cow (Q6) as predicted by the length of the book closest to the survey respondent (Q5). 
```{r LinearRegression}

Model1 <- lm(Q6 ~ Q5, data = w1df) #<<

Model1
```
---
# Let's clean up our language...

In this situation, Q6 (the Estimated weight of the cow) is the Outcome Variable and Q5 (the number of pages in the closest book) is the Predictor variable.

Outcome Variable is often referred to as independent variable

Predictor Variable is often referred to as the dependent variable. 
 
I hate this language. Predictor and Outcome are much more descriptive. 

---
# Let's figure out what this means

```{r ModelInterp}

Model1

```
**Intercept** = if the closest book had 0 pages in it, this would be the estimated weight of the cow that we predict. 

**Q6 coefficient** = If we increase the number of pages in the book, the value of the coeffient tells us how much we would increase the estimated weight of the cow by 


---
# Let's check out those residuals...
.pull-left[
```{r ScatterPlotWithResiduals, echo=FALSE }

mod_df<-augment(Model1)


mod_df %>%
  ggplot(aes(x = Q5, y = Q6)) +
  geom_point() + 
  theme_minimal() + 
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = Q5, yend= .fitted), color = "red", size = 0.2) +
  labs(x = "Pages", 
       y = "Weight of Cow",
       title = "Kick Ass Scatterplot!")

```

]
.pull-right[




```{r ResidualCheck2}

augment(Model1) %>%
  select(Q5:.resid) 
```
]

---
# Let's see how much better the model fits

```{r ResidualCheck3}

augment(Model1) %>%
  select(Q5:.resid) %>%
  mutate(SqResid = .resid^2) #<<
```

---
# Let's see how much better the model fits

```{r ResidualCheck4}

augment(Model1) %>%
  select(Q5:.resid) %>%
  mutate(SqResid = .resid^2) %>%
  tally(SqResid)
```

---
# Let's take a look at R<sup>2</sup>


**R<sup>2</sup>** is the measure of how well the model fits

$R^{2} = 1 - \frac{\sum_{i} (Y_{Pred} - Y_{i})^2}{\sum_{i} (Y_{i}-\overline{Y})^2}$


```{r  CalcRSq}

augment(Model1) %>%
  select(Q6:.resid) %>%
  mutate(SqResid = .resid^2) %>%
  mutate(SqDiffMean = (Q6 - mean(Q6))^2) -> resid_df

glimpse(resid_df)

```

---
# Let's take a look at R<sup>2</sup>


$R^{2} = 1 - \frac{\sum_{i} (Y_{Pred} - Y_{i})^2}{\sum_{i} (Y_{i}-\overline{Y})^2}$


```{r  CalcRSq2}

SquaredResiduals = sum(resid_df$SqResid)
SquaredResiduals

SquaredDiffMean = sum(resid_df$SqDiffMean)
SquaredDiffMean

r2 = 1-(SquaredResiduals/SquaredDiffMean)
r2
```

---
# Let's compare to the R<sup>2<\sup>

```{r ModelSummary}
SquaredResiduals
SquaredDiffMean
r2

summary(Model1)

```



---
# Let's recap
We can...
- Fit a regression model to two continuous variables
- Calculate the residuals
- Calculate the squared residuals
- Use the residuals to calculate the R<sup>2</sup>

    
    