---
title: "Workshop 1"
draft: false
date: "2020-07-21T00:00:00+01:00"
output: html_document
menu:
  RForSNA:
    parent: Workshops
    weight: 2
linktitle: "Workshop 1"
type: docs
weight: 2
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

In order to prepare for workshop one - I have five things that I need you to do:

1. Install R and RStudio on your computer. 
    a. R can be downloaded from a link on this website <https://www.r-project.org/>
    b. RStudio can be downloaded here: <https://rstudio.com/products/rstudio/download/>
2. Set up a series of folders (on your computer or Dropbox / Box.com- don't use Google Drive or OneDrive):
    a. The top level folder should be called "RForSNA"
    b. Inside the folder "RForSNA" there should be two folders called "data" and "plots"
3. I will send you a survey on Survey Monkey. I need you to answer the survey
4. Before the first workshop, I will email you a few csv files, please download each and save in the folder titled "data"
5. Let me know by email if you have trouble with steps 1-4.

The slides for Workshop #1 can be found here:
```{r WS1Slides}
knitr::include_url('https://ericbrewe.com/slides/rforsna/rforsna_ws1.html#1')
```

